Global Climate and Atmospheric Composition of the Ultra-Hot Jupiter WASP-103\lowercase{b} from \HST\ and \Spitzer\ Phase Curve Observations

### 2018-04-18 ApJ_submit_v2/response.txt

****

14. The authors close section 5.2.1 with a list of caveats about the error bars on this section and how the inferred [M/H] and [C/O] may be heavily skewed by the one 4.5 micron data point (from phase 0.5? or other bin? see previous comment). These caveats need to be more prominent considering how serious they may be to the interpretation. I think they need to be present in the abstract, and also they need to be bulleted here in the relevant section so that they are easy to catch for those that skim the paper instead of read it in depth.

#

We have reformatted section 5.2.1 to include the caveats as a bulleted list.  Since the caveats are now stated very prominently in 5.2.1 and in the summary bullet point in Section 8, we are confident that readers will not miss them. We did not add these points to the abstract because we had to cut it substantially to satisfy the length limit, and we feel the impact of the paper will be greater if we focus scientific results in the abstract and leave the caveats in the main text.

@@

17. The discussion of the magnetic field strength needed to produce a drag that sort of makes the GCM model work the best is very wishy washy. I understand that the authors are using order-of-magnitude expressions (as they state) but I think the magnetic field discussion is oversimplified and the conclusions are overstated. It also neglects to mention previous work attempting to directly measure/detect the effects of magnetic fields from planets. If the authors wish to discuss the magnetic fields, some mention of these previous works should be done. Significant mention about the shape of the magnetic field should also be prevalent. Furthermore, the planet is very close in to its star, and so there would likely be a complicated magnetic field structure due to the presence of the stellar magnetic field in the vicinity. This is not addressed or touched on when the authors assume the magnetic field is perpendicular to the flow of material.

If the authors want to keep a discussion of the magnetic field in the paper, this section should be significantly expanded and the caveats and assumptions more prominently listed. Otherwise, casual readers might over interpret this section to imply that a magnetic field is detected or that a magnetic field with a strength nice and similar to Jupiter's is the only possible explanation for the observable properties of the system with regards to its phase curve. It is the referee's preference however, that the magnetic field discussion is mostly scaled down or eliminated.

#

We expanded the discussion of magnetic fields according to the referee's suggestions. We think it is important to keep this section of the paper because it opens up discussion about a new way of measuring planet magnetic fields (studying their effects on the atmospheric dynamics, in contrast to previous work, which focused on star-planet interactions or planetary radio emission).  We put this discussion in a new subsection (6.1), added references to previous work, and mentioned the suggested caveats. We removed the specific mention of the magnetic field estimate from the abstract.

### 2018-04-18 ApJ_submit_v2/ms.tex
shorten abstract

%	We present thermal phase curve measurements for the hot Jupiter WASP-103b observed with \emph{Hubble}/WFC3 and \emph{Spitzer}/IRAC. The dayside planet-to-star flux ratios are $0.151\pm0.015\%$, $0.446\pm0.038\%$, and $0.569\pm0.014\%$ in bandpasses centered at $1.4$, $3.6$, and $4.5\mu$m, respectively.  The phase curves have large amplitudes and negligible hotspot offsets, indicative of poor heat redistribution to the nightside.  We fit the phase variation with a range of climate maps and find that a spherical harmonics model generally provides the best fit. The phase-resolved spectra are consistent with blackbodies in the WFC3 bandpass, with brightness temperatures ranging from $1880\pm40$ K on the nightside to $2930 \pm 40$ K on the dayside. The dayside spectrum has a significantly higher brightness temperature in the \Spitzer\ bands, which we attribute to CO emission and a thermal inversion.  The inversion is not present on the nightside. The transmission spectrum is featureless, but consistent with the small transit depth variation expected from cloud-free models. Using the dayside and nightside emission spectra, we retrieved the atmospheric composition with a 1D self-consistent modeling technique and found the atmosphere is moderately metal-enriched ($\mathrm{[M/H]} =  23^{+29}_{-13}\times$ solar) and the carbon-to-oxygen ratio is below 0.9 at $3\,\sigma$ confidence. In contrast to cooler hot Jupiters, we do not detect significant water absorption, which we attribute to partial H$_2$O dissociation and additional opacity from H- and trace amounts of hydrides/oxides.  We compare the phase curves to 3D general circulation models and find that atmospheric drag with a timescale of $10^4$ s is the best match to the data, which can be produced with a simple model of the magnetic field with strength $\sim1$ Gauss.  We also compare the WASP-103b spectra to brown dwarfs and young directly imaged companions and find these more massive and self-luminous objects have significantly larger water features, indicating that surface gravity and irradiation environment play an important role in shaping the spectra of hot Jupiters. These results highlight the 3D structure of exoplanet atmospheres and illustrate the importance of phase curve observations for understanding their complex chemistry and physics.
	We present thermal phase curve measurements for the hot Jupiter WASP-103b observed with \emph{Hubble}/WFC3 and \emph{Spitzer}/IRAC.  The dayside planet-to-star flux ratios are $0.151\pm0.015\%$, $0.446\pm0.038\%$, and $0.569\pm0.014\%$ in bandpasses centered at $1.4$, $3.6$, and $4.5\mu$m, respectively.  The phase curves have large amplitudes and negligible hotspot offsets, indicative of poor heat redistribution to the nightside.  We fit the phase variation with a range of climate maps and find that a spherical harmonics model generally provides the best fit. The phase-resolved spectra are consistent with blackbodies in the WFC3 bandpass, with brightness temperatures ranging from $1880\pm40$ K on the nightside to $2930 \pm 40$ K on the dayside. The dayside spectrum has a significantly higher brightness temperature in the \Spitzer\ bands, which we attribute to CO emission and a thermal inversion.  The inversion is not present on the nightside. The transmission spectrum is featureless, but consistent with the small transit depth variation expected from cloud-free models. Using the dayside and nightside emission spectra, we retrieved the atmospheric composition with a 1D self-consistent modeling technique and found the atmosphere is moderately metal-enriched ($\mathrm{[M/H]} = 23^{+29}_{-13}\times$ solar) and the carbon-to-oxygen ratio is below 0.9 at $3\,\sigma$ confidence. In contrast to cooler hot Jupiters, we do not detect significant water absorption, which we attribute to partial H$_2$O dissociation and additional opacity from H- and trace amounts of hydrides/oxides.  We compare the phase curves to 3D general circulation models and find that magnetic drag effects are needed match the data; \added{a back-of-the-envelope calculation suggests the strength of the magnetic field is $\sim1$ Gauss.}  We also compare the WASP-103b spectra to brown dwarfs and young directly imaged companions and find these more massive and self-luminous objects have significantly larger water features, indicating that surface gravity and irradiation environment play an important role in shaping the spectra of hot Jupiters. These results highlight the 3D structure of exoplanet atmospheres and illustrate the importance of phase curve observations for understanding their complex chemistry and physics.

@@

\subsection{Transmission Spectrum}
Each phase curve observation includes a transit of WASP-103b. To measure the
wavelength-dependent transit depths (the transmission spectrum), we select a
subset of each phase curve. The subset includes the transit and additional
baseline on either side, such that the total light curve has twice the duration
of the transit. Over this short duration, there is negligible curvature in the
light curve due to the planet's thermal phase variation. We fit the data with a
transit model, which has free parameters for the planet-to-star radius and a
linear limb darkening parameter. The orbital parameters (inclination, $a/R_s$,
period, and time of central transit) were fixed at previously published values
listed in \S\,\ref{sec:fits}. We fit for the instrument systematics using the
same model as for the phase curve fits, except that we modeled the visit-long
systematic as a linear trend in time (which is sufficient for the shorter
duration). The advantage of fitting the transits separately from the full phase
curves is that the resultant transit depths are not dependent on how the phase
variation is modeled. The transit light curve fits are shown in
Figure\,\ref{fig:transitlc}.

\begin{figure}
\includegraphics[width = 0.5\textwidth]{fig9.pdf}
\caption{Best fit transit light curves for WASP-103b (lines), compared to binned data (black points). From top-to-bottom, we show the band-integrated WFC3 light curve, \Spitzer\ 3.6, and $4.5\,\mu$m.}
\label{fig:transitlc}
\end{figure}

We show the measured transmission spectrum in Figure\,\ref{fig:tspec}.  The spectrum is biased by flux contamination from the planet's nightside; to correct for nightside emission, we estimate the average nightside planet-to-star flux ratio and subtract it from the transit depth. We calculate planet-to-star flux using a NextGen spectrum for the star (interpolated to $T_\mathrm{eff} = 6110$ K), a blackbody for the planet, and a planet-to-star radius ratio of 0.1146. We assume a nightside temperature of 1700 K, which is near the median of the nightside temperatures estimated from the phase variation models (Table\,\ref{table:models}). We also show the uncorrected transit depths. The corrected and uncorrected transit depths are listed in Table\,\ref{table:tspec}. For the uncorrected data, there is an offset between the \HST\ and \Spitzer\ data of more than five atmospheric scale heights.  The uncorrected spectrum is inconsistent with a flat line at $5.3\,\sigma$ confidence, whereas the nightside-corrected spectrum is consistent within $1\,\sigma$.  The corrected spectrum is also consistent with predictions from the $\tau_\mathrm{drag4}$ GCM, which shows water features in the WFC3 bandpass. Future observations with higher precision could reveal these features. %The mean transit depths are $1.3431 \pm 0.0029$,  $1.3360 \pm 0.0029$, and $1.3078 \pm 0.0029\%$ for assumed nightside temperatures of 0, 1500, and 2000 K.  Of these, only the 2000 K model yields a transit depth consistent with the measurement of \cite{lendl17} at shorter wavelengths where nightside contamination is negligible ($1.3092 \pm 0.0114$\%). FIXME: could we see water at higher precision?

\begin{figure}
\includegraphics[width = 0.5\textwidth]{fig10.pdf}
\label{fig:tspec}
\caption{The transmission spectrum of WASP-103b, corrected for planet nightside emission at 1700 K (blue points) and uncorrected (red points). The dark gray line is the model transmission spectrum from the $\tau_\mathrm{drag4}$ GCM, with squares indicating the model binned over the \Spitzer\ bandpasses. Atmospheric scale height $H$ is shown on the right y-axis, where $H = 5.5\times10^6$ m (assuming a mean molecular weight of 2.3 atomic mass units, surface gravity $g = 15.9$ m/s$^2$, and a temperature $T = 2410$ K).}
\end{figure}

@@

\subsection{Constraints on the Planet's Magnetic Field}
\added{We showed in this section that the small observed hotspot offset in the phase curves is best fit by a GCM that includes magnetic drag with a timescale $\tau_\mathrm{drag} = 10^4$\,s. This observation gives rise to the question: what magnetic field strength on the planet can produce drag with this timescale?  Previous efforts to characterize exoplanet magnetic fields have mainly focused on magnetic interaction between the planet and its host star \citep[e.g.][and references therein]{wright15} and planetary radio emission \citep{griessmeier15}.  A complementary approach is to study the effect of the magnetic field on the planet's atmospheric dynamics.}

>>>>

	%We present thermal phase curve measurements for the hot Jupiter WASP-103b observed with \emph{Hubble}/WFC3 and \emph{Spitzer}/IRAC.  The dayside planet-to-star flux ratios are $0.151\pm0.015\%$, $0.446\pm0.038\%$, and $0.569\pm0.014\%$ in bandpasses centered at $1.4$, $3.6$, and $4.5\mu$m, respectively.  The phase curves have large amplitudes and negligible hotspot offsets, indicative of poor heat redistribution to the nightside.  We fit the phase variation with a range of climate maps and find that a spherical harmonics model generally provides the best fit. The phase-resolved spectra are consistent with blackbodies in the WFC3 bandpass, with brightness temperatures ranging from $1880\pm40$ K on the nightside to $2930 \pm 40$ K on the dayside. The dayside spectrum has a significantly higher brightness temperature in the \Spitzer\ bands, which we attribute to CO emission and a thermal inversion.  The inversion is not present on the nightside. The transmission spectrum is featureless, but consistent with the small transit depth variation expected from cloud-free models. Using the dayside and nightside emission spectra, we retrieved the atmospheric composition with a 1D self-consistent modeling technique and found the atmosphere is moderately metal-enriched ($\mathrm{[M/H]} = 23^{+29}_{-13}\times$ solar) and the carbon-to-oxygen ratio is below 0.9 at $3\,\sigma$ confidence. In contrast to cooler hot Jupiters, we do not detect significant water absorption, which we attribute to partial H$_2$O dissociation and additional opacity from H- and trace amounts of hydrides/oxides.  We compare the phase curves to 3D general circulation models and find that magnetic drag effects are needed to match the data; \added{a back-of-the-envelope calculation suggests the strength of the magnetic field is $\sim1$ Gauss.}  We also compare the WASP-103b spectra to brown dwarfs and young directly imaged companions and find these more massive and self-luminous objects have significantly larger water features, indicating that surface gravity and irradiation environment play an important role in shaping the spectra of hot Jupiters. These results highlight the 3D structure of exoplanet atmospheres and illustrate the importance of phase curve observations for understanding their complex chemistry and physics.
We present thermal phase curve measurements for the hot Jupiter WASP-103b observed with \emph{Hubble}/WFC3 and \emph{Spitzer}/IRAC.  The phase curves have large amplitudes and negligible hotspot offsets, indicative of poor heat redistribution to the nightside.  We fit the phase variation with a range of climate maps and find that a spherical harmonics model generally provides the best fit. The phase-resolved spectra are consistent with blackbodies in the WFC3 bandpass, with brightness temperatures ranging from $1880\pm40$ K on the nightside to $2930 \pm 40$ K on the dayside. The dayside spectrum has a significantly higher brightness temperature in the \Spitzer\ bands, likely due to CO emission and a thermal inversion.  The inversion is not present on the nightside. We retrieved the atmospheric composition and found the composition is moderately metal-enriched ($\mathrm{[M/H]} = 23^{+29}_{-13}\times$ solar) and the carbon-to-oxygen ratio is below 0.9 at $3\,\sigma$ confidence. In contrast to cooler hot Jupiters, we do not detect spectral features from water, which we attribute to partial H$_2$O dissociation.  We compare the phase curves to 3D general circulation models and find magnetic drag effects are needed to match the data.  We also compare the WASP-103b spectra to brown dwarfs and young directly imaged companions and find these objects have significantly larger water features, indicating that surface gravity and irradiation environment play an important role in shaping the spectra of hot Jupiters. These results highlight the 3D structure of exoplanet atmospheres and illustrate the importance of phase curve observations for understanding their complex chemistry and physics.

@@

\subsection{Transmission Spectrum}
Each phase curve observation includes a transit of WASP-103b. To measure the
wavelength-dependent transit depths (the transmission spectrum), we select a
subset of each phase curve. The subset includes the transit and additional
baseline on either side, such that the total light curve has twice the duration
of the transit. Over this short duration, there is negligible curvature in the
light curve due to the planet's thermal phase variation. We fit the data with a
transit model, which has free parameters for the planet-to-star radius and a
linear limb darkening parameter. The orbital parameters (inclination, $a/R_s$,
period, and time of central transit) were fixed at previously published values
listed in \S\,\ref{sec:fits}. We fit for the instrument systematics using the
same model as for the phase curve fits, except that we modeled the visit-long
systematic as a linear trend in time (which is sufficient for the shorter
duration). The advantage of fitting the transits separately from the full phase
curves is that the resultant transit depths are not dependent on how the phase
variation is modeled. The transit light curve fits are shown in
Figure\,\ref{fig:transit_light_curves}.

\begin{figure}
\includegraphics[width = 0.5\textwidth]{fig9.pdf}
\caption{Best fit transit light curves for WASP-103b (lines), compared to binned data (black points). From top-to-bottom, we show the band-integrated WFC3 light curve, \Spitzer\ 3.6, and $4.5\,\mu$m.}
\label{fig:transit_light_curves}
\end{figure}

We show the measured transmission spectrum in Figure\,\ref{fig:tspectrum}.  The spectrum is biased by flux contamination from the planet's nightside; to correct for nightside emission, we estimate the average nightside planet-to-star flux ratio and subtract it from the transit depth. We calculate planet-to-star flux using a NextGen spectrum for the star (interpolated to $T_\mathrm{eff} = 6110$ K), a blackbody for the planet, and a planet-to-star radius ratio of 0.1146. We assume a nightside temperature of 1700 K, which is near the median of the nightside temperatures estimated from the phase variation models (Table\,\ref{table:models}). We also show the uncorrected transit depths. The corrected and uncorrected transit depths are listed in Table\,\ref{table:tspec}. For the uncorrected data, there is an offset between the \HST\ and \Spitzer\ data of more than five atmospheric scale heights.  The uncorrected spectrum is inconsistent with a flat line at $5.3\,\sigma$ confidence, whereas the nightside-corrected spectrum is consistent within $1\,\sigma$.  The corrected spectrum is also consistent with predictions from the $\tau_\mathrm{drag4}$ GCM, which shows water features in the WFC3 bandpass. Future observations with higher precision could reveal these features. %The mean transit depths are $1.3431 \pm 0.0029$,  $1.3360 \pm 0.0029$, and $1.3078 \pm 0.0029\%$ for assumed nightside temperatures of 0, 1500, and 2000 K.  Of these, only the 2000 K model yields a transit depth consistent with the measurement of \cite{lendl17} at shorter wavelengths where nightside contamination is negligible ($1.3092 \pm 0.0114$\%). FIXME: could we see water at higher precision?

\begin{figure}
\includegraphics[width = 0.5\textwidth]{fig10.pdf}
\caption{The transmission spectrum of WASP-103b, corrected for planet nightside emission at 1700 K (blue points) and uncorrected (red points). The dark gray line is the model transmission spectrum from the $\tau_\mathrm{drag4}$ GCM, with squares indicating the model binned over the \Spitzer\ bandpasses. Atmospheric scale height $H$ is shown on the right y-axis, where $H = 5.5\times10^6$ m (assuming a mean molecular weight of 2.3 atomic mass units, surface gravity $g = 15.9$ m/s$^2$, and a temperature $T = 2410$ K).}
\label{fig:tspectrum}
\end{figure}

@@

\subsection{Constraints on the Planet's Magnetic Field}
\added{We show in this section that the small observed hotspot offset in the phase curves is best fit by a GCM that includes Rayleigh drag with a timescale $\tau_\mathrm{drag} = 10^4$\,s. This observation gives rise to the question: what magnetic field strength on the planet can produce drag with this timescale?  Previous efforts to characterize exoplanet magnetic fields have mainly focused on magnetic interaction between the planet and its host star \citep[e.g.][and references therein]{wright15} and planetary radio emission \citep{griessmeier15}.  A complementary approach is to study the effect of the magnetic field on the planet's atmospheric dynamics.}
